---
title: "lab_02"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---
## Akira Kyles, 9/17/25, Jour772

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
install.packages('tidyverse')
install.packages('lubridate')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **A1.**
##There have been 32 departments that offered courses with "climate" in the title and at least one seat. Atmospheric and Oceanic Science department has offered the most courses with "climate" in the title. 

```{r}
umd_courses |> 
  filter(str_detect(title, "Climate") & seats >= 1) |> 
  group_by(department) |> 
  summarise(
    count_classes = n()
  ) |>
  arrange (desc(count_classes))
```

**Q2.** Of the departments listed in A1, which one is the most interesting to you, and why? **A2.**
## Anthropology is the most interesting to me because I find the study of humanity fascinating. There are so many aspects of life that makes the human experience so unique. 

**Q3** Which departments offered the most courses in the Fall 2024 term? And which departments in that term have at least one syllabus posted? Answer both of those questions and then compare the top five departments - how are they different?

**A3.**
#The Business and Management department, School of Music department, and Economics offered the most courses in the Fall 2024 term. Economics, Business and Managment and Philosophy were among 134  departments that had at least one syllabus posted. Departments with higher syllabus counts have lower course counts. In the first five departments, at least three of them show up in the top five for both results. The first included the School of music department and Information studies with the the most courses offered but when compared by syllabus count, they weren't in the top five anymore, instead Philosophy and First-Year Innovation & Research Experience were. 


```{r}
umd_courses |> 
  filter(str_detect(term, "202408") & syllabus_count >= 1) |> 
 group_by(department) |> 
  summarise(
    count_classes = n()
  ) |>
  arrange (desc(count_classes))
```
```{r}
umd_courses |> 
  filter(str_detect(term, "202408") & syllabus_count >= 1) |> 
  group_by(department) |> 
  summarise(total_amount = sum(syllabus_count)
  ) |>
  arrange (desc(total_amount))
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as a new dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name, and we're not looking in payee_name). Write a sentence describing the pattern of spending over time. **A4.**
#From the data I can see that spending continuously increased over time.  
```{r}
maryland_expenses <- maryland_expenses |> mutate(expenditure_date=mdy(expenditure_date))
```

```{r}
head(maryland_expenses)
```
```{r}
maryland_expenses |>
  mutate(month = floor_date(expenditure_date, "month")) |>
  group_by(month) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount))
```


**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for all others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **A5**
##There was more money spent in-state compared to out of state. There was $77,723,146 spent in-state and $33,164,994 spent out of state. 

```{r}
spending_location <- maryland_expenses |> 
  mutate(
    state = case_when(
      str_detect(address, " Maryland ") ~ "In-state",
        str_detect(address, " California ") ~ "Out of State",
        str_detect(address, " Washington ") ~ "Out of State",
        str_detect(address, " Louisiana ") ~ "Out of State",
        str_detect(address, " Florida ") ~ "Out of State",
        str_detect(address, " North Carolina ") ~ "Out of State",
        str_detect(address, " Massachusetts ") ~ "Out of State",
        str_detect(address, " West Virginia ") ~ "Out of State",
        str_detect(address, " Virginia ") ~ "Out of State",
        .default = NA
      )
  )
```

```{r}
spending_location |> 
  group_by(state) |> 
  summarize(total = sum(amount)) |> 
  arrange(desc(total))
```
**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **A6.**
##What course had the most seats in the Spring 2024 semester?
##Doctoral Dissertation Research had the most seats in the Spring 2024 semester. 
```{r}
umd_courses |> 
  filter(str_detect(term, "202405") & seats) |> 
  group_by(id, title) |> 
  summarise(total_amount = sum(seats)
  ) |>
  arrange (desc(total_amount))
```
