---
title: "lab_03"
author: "Derek Willis/ Rob Wells"
date: "2025-09-24"
output: html_document
---

###Akira Kyles, 9/24/2025, Jour 772
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TASK: Load libraries

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(janitor)
# Load the tidyverse

# Load janitor
```

## TASK: Load data and clean up the column names

Load the two datasets we worked with in the pre_lab that are in the data folder: the Maryland payments to vendors and Maryland grants and loans AND, using janitor, change the column names so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
# Load the data
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv") |> 
  clean_names()

# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv") |> 
  clean_names()
```
## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed in the payments data. 

Look at & search the records in the dataframe to start your search, starting with a broad search and then narrowing it down. Then write code to find rows where the vendor name contains some version of the county's school system.

HINT: you should use filter() just once, and inside your filter you want to look for the variations one at a time. Think "this version OR that version."

When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?
### I found 64 records. 

A1:

```{r}
pg_schools <- payments |> 
  filter(
    str_detect(vendor_name,"PRINCE GEORGES COUNTY PUBLIC SCHOOLS")|  
 str_detect(vendor_name,"PRINCE GEORGES PUBLIC SCHOOLS")|
  str_detect(vendor_name,"PG COUNTY PUBLIC SCHOOLS")|
 str_detect(vendor_name,"PRINCE GEORGE'S COUNTY PUBLIC SCHOOLS")|
   str_detect(vendor_name,"PRINCE GEORGES COUNTY PUBLIC SCHOOL")|
  str_detect(vendor_name,"PRINCE GEORGES COUNTY SCHOOL"))
  
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? Are there any aspects of the results that stand out to you?

A2:
### The fiscal year 2021 had the largest total.I notice there's a gap in the results for 2022 to 2024. I also notice that 2025 is almost over and it has the lowest amount for the fiscal year.  

```{r}
pg_schools |>
  group_by(fiscal_year) |> 
  summarize(total_amount = sum(amount)) |> 
  arrange(desc(total_amount))
  
```

Q3. This question is in two parts. 

Part 1: Code: You're interested in examining the vendors paid for capital projects at the University of Maryland. Write code that does the following: 
--Filters payments data to capital projects at the University of Maryland
--Produce a list of vendors and the total amount they received. 

Part 2: Provide a written answer, not code.

Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? 
--Provide at least three examples of vendor names that are similar and should be combined into one. 
--Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

###There are multiple copies of vendor names and multiple variations of vendor names so we would have to clean the data so it pools together data from each variation of vendor names.
### Vendor EX.1: A MORTON THOMAS & ASSOC INC			
###              A MORTON THOMAS & ASSOCIATES INC
### Vendor EX.2: AEROSOL MONITORING & ANALYSIS INC	
###              AERSOL MONITORING & ANALYSIS INC
### Vendor EX.3: AMERICAN DESIGN ASSOCIATES			
###              AMERICAN DESIGN ASSOCIATES INC

### For the multiple variations, I would use filter and str_detect to gather all the variations and I would use clean_dup to pool all the duplicates together. 

A3.

```{r}
payments |> 
  filter(
    str_detect(agency_name,'CAPITAL PROJECTS- UNIVERSITY OF MARYLAND')) |> 
  group_by(vendor_name, amount) |> 
summarize(total_amount = sum(amount))

```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. 

Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. 
Then create a total amount of money for each fiscal year. 

Write a sentence summarizing how stem cell research funding has changed over time. Be specific with the totals.

A4:
##Grants for stem cell research were at their highest over a decade ago and only recently, in 2023, exceeded $10 million in funding. 

```{r}
stem_cells <- md_grants_loans |> 
  filter(
    str_detect(description,"STEM CELL RESEARCH GRANT")| 
    str_detect(description,"Stem Cell Research Grant, Pre-Commercialization Research")|
    str_detect(description,"Stem Cell Research Grant"))
```

```{r}
stem_cells |> 
  group_by(fiscal_year) |> 
summarize(total_amount = sum(amount)) |> 
arrange(desc(fiscal_year))

```

Q5: Using the Maryland grants and loan data, find records where the zip code is the one where the University of Maryland-College Park is located in and produce a list of grantees from that zip code, along with the total amount they received. Write a paragraph explaining how many grantees are there, what you would need to do to clean the results so that you had one record per grantee and which grantees are most interesting to you?
###There are 18 grantees.In order to clean the data so there's one record per grantee, I would need to use the filter function in order to pool the different variations of UMD to include UMD departments getting funding. Although, it's not that much in the grand scheme of departmental funding, I still found it intesting that the arts department received six figures in funding.  
A5:

```{r}
cp_zips <- md_grants_loans |>
  filter(
    str_detect(zip_code,"20742")|
    str_detect(zip_code,"207423141")
  ) 

```
```{r}
cp_zips |> 
  group_by(grantee,zip_code) |> 
  summarise(total_amount = sum(amount)) |> 
  arrange(desc(total_amount))
```
